---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE, warning = FALSE,
  comment = "#>", fig.align = "center",
  fig.path = "man/figures/README-", dpi = 120
)
```
# mists

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/mists)](https://cran.r-project.org/package=mists)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The **mists** package provides a suite of 1d, 2d, and visual tools for exploring and polishing missing values residing in temporal data. The primary focus of temporal missings is to look at the runs of `NA`, and the association with other variables.

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("earowang/mists")
```

## Get started with `na_rle()` and `list_of_na_rle()`

The `na_rle()` gives a sparser representation for runs of missings, a special type of run length encoding (`rle`).

```{r na-rle}
library(mists)
(x <- na_rle(c(1, NA, NA, 4:7, NA, NA, 10:15, NA)))
(y <- na_rle(c(10, NA, NA, NA, 6:3, NA, 1)))
```

It returns a named list of `lengths` and `values` (the starting indices) of `NA` runs. Set operations can be applied: (1) `intersect(x, y)`, (2) `union(x, y)`, (3) `setdiff(x, y)`.

The `list_of_na_rle()` makes it easier to work with tibbles. 

```{r list-of-na-rle}
library(dplyr)
na_runs_wind <- nycflights13::weather %>% 
  group_by(origin) %>% 
  summarise_at(vars(contains("wind")), ~ list_of_na_rle(., time_hour))
na_runs_wind
```

## Range plots and extended spinoplots

How are those missings distributed and associated with others? 

```{r autoplot, fig.height = 2}
autoplot(na_runs_wind$wind_dir, na_runs_wind$origin, shape = 4)
```

```{r spinoplot, fig.width = 4, fig.height = 3}
na_rle_spinoplot(na_runs_wind$wind_dir[[1]], na_runs_wind$wind_gust[[1]])
```

## Data polishing

Data polishing isn't data cleaning.

## Related work

* [imputeTS](http://steffenmoritz.github.io/imputeTS/): Time Series Missing Value Imputation
* [VIM](https://github.com/statistikat/VIM): Visualization and Imputation of Missing Values
* [naniar](http://naniar.njtierney.com): Data Structures, Summaries, and Visualisations for Missing Data
