---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, message = FALSE,
  comment = "#>", fig.align = "center",
  fig.path = "man/figures/README-", dpi = 120
)
```
# mists

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/earowang/mists.svg?branch=master)](https://travis-ci.org/earowang/mists)
[![Codecov test coverage](https://codecov.io/gh/earowang/mists/branch/master/graph/badge.svg)](https://codecov.io/gh/earowang/mists?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/mists)](https://cran.r-project.org/package=mists)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->

The **mists** package provides a suite of 1d, 2d, and visual tools for exploring and polishing missing values residing in temporal data. The primary focus of the package is to look at the runs of `NA` and the association with other variables, and to formulate missing data polishing strategies.

## Installation

Install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("earowang/mists")
```

## Get started with `na_rle()` and `list_of_na_rle()`

The `na_rle()` gives a sparser representation for runs of missings, a special type of run length encoding (`rle`).

```{r na-rle}
library(mists)
(x <- na_rle(c(1, NA, NA, 4:7, NA, NA, 10:15, NA)))
(y <- na_rle(c(10, NA, NA, NA, 6:3, NA, 1)))
```

It returns a named list of `lengths` and `values` (the starting indices) of `NA` runs. Set operations can be applied: (1) `intersect(x, y)`, (2) `union(x, y)`, (3) `setdiff(x, y)`.

The `list_of_na_rle()` makes it easier to work with tibbles. 

```{r list-of-na-rle}
library(dplyr)
na_runs_wind <- nycflights13::weather %>% 
  group_by(origin) %>% 
  summarise_at(
    vars(contains("wind")), 
    ~ list_of_na_rle(., index_by = time_hour)
  )
na_runs_wind
```

## Range plots and extended spinoplots

How do those missings distribute and associate with others? The range plot below shows many of single `NA` runs occurs to the `wind_dir` variable over the year of 2013.

```{r rangeplot, fig.height = 2}
na_runs_wind %>% 
  na_rle_rangeplot(x = wind_dir, y = origin, shape = 4)
```

The spinoplot has been extended to examine the association of missing runs with a second variable through temporal overlapping. A large amount of missings in `wind_dir` intersects with `wind_gust` missings at all three places.

```{r spinoplot, fig.height = 3}
na_runs_wind %>% 
  na_rle_spinoplot(x = wind_dir, y = wind_gust, facets = origin)
```

## Missing data polishing

Too many missings spread across variables and observations like the `wdi` dataset (world development indicators)? Every observation and measured variable contains `NA`, and almost half of the data goes missing. We would end up with no data if using listwise deletion.

```{r wdi-na-metrics, echo = FALSE}
report_na_metrics <- function(data) {
  tibble(
    prop_overall_na = mean(is.na(data[, -(1:2)])),
    prop_cols_na = mean(apply(data[, -(1:2)], 2, anyNA)),
    prop_rows_na = mean(apply(data[, -(1:2)], 1, anyNA)),
    data_ncols = NCOL(data),
    data_nrows = NROW(data)
  )
}
report_na_metrics(wdi)
```

Polishing columns and rows iteratively sweeps out chunks of `NA` but leaves some of them in. The index should be kept intact by slicing the ends only due to the temporal ordering. The `na_polish_auto()` automates this polishing process by minimising *the proportion of overall missings* weighted by *the proportion of removed observations*. It will go through the following passes several times: `na_polish_measures()`, `na_polish_key()`, `na_polish_index()` until a tolerance value close or equal to zero. The `na_polish_*()` family expects a [tsibble](http://tsibble.tidyverts.org), because polishing starts with tidy and clean data.

```{r polish}
wdi_ts <- wdi %>% 
  tsibble::as_tsibble(key = country_code, index = year)
wdi_after <- na_polish_auto(wdi_ts, cutoff = .8, quiet = TRUE)
na_polish_metrics(wdi_ts, wdi_after)
```

## Related work

* [imputeTS](http://steffenmoritz.github.io/imputeTS/): Time Series Missing Value Imputation
* [VIM](https://github.com/statistikat/VIM): Visualization and Imputation of Missing Values
* [naniar](http://naniar.njtierney.com): Data Structures, Summaries, and Visualisations for Missing Data
